---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(tidymodels)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Food nutrients] This problem will use PCA to provide a low-dimensional view
of a 14-dimensional nutritional facts
[dataset](https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv).
The data were originally curated by the USDA and are regularly used in
[visualization studies](http://bl.ocks.org/syntagmatic/raw/3150059/).
    
```{r}
nutrients <- read_csv("https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv")
```
    
a. Define a tidymodels `recipe` that normalizes all nutrient features and The
code below defines a recipe that first normalizes the predictors and then
performs PCA.

```{r}
# pca_recipe <- 
```

b. Visualize the top 6 principal components. What types of food do you expect to
have low or high values for PC1 or PC2?

```{r, fig.width = 10, fig.height = 8}
```

c. Compute the average value of PC2 within each category of the `group` column.
Give the names of the groups sorted by this average.

```{r}
```

d. Visualize the scores of each food item with respect to the first two
principal components. Facet the visualization according to the `group` column,
and sort the facets according to the results of part (c). How does the result
compare with your guess from part (b)?

```{r}
```

