---
title: "Demo 2 - 1"
output: rmdformats::readthedown
date: "`r Sys.Date()`"
css: custom.css
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

[Plant Growth Experiment] This problem will give you practice with tidying a
dataset so that it can be easily visualized. The data describe the height of
several plants measured every 7 days. The plants have been treated with
different amounts of a growth stimulant. The first few rows are printed below --
`height.x` denotes the height of the plant on day `x`.

```{r}
library(tidyverse)
plants <- read_csv("https://uwmadison.box.com/shared/static/qg9gwk2ldjdtcmmmiropcunf34ddonya.csv")
plants
```

a. Propose an alternative arrangement of rows and columns that conforms to the
tidy data principle.

# Section B

b. Implement your proposed arrangement from part (a).

```{r}
  #pivot_longer(c("height.0", "height.7", "height.14", "height.21"))
  #pivot_longer(height.0:height.21)

plants_long <- plants |>
  pivot_longer(starts_with("height"), names_to = "time", values_to = "height") |>
  separate("time", c("dummy", "time"), convert = TRUE)

#plants_long <- plants_long |>
#  mutate(
#    time = str_extract(time, "[0-9]+"),
#    time = as.integer(time)
#    )
```


c. Using the dataset from (b), design and implement a visualization showing the
growth of the plants over time according to different treatments.

```{r}
my_theme <- theme_bw() +
  theme(
    axis.title = element_text(size = 16),
    legend.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    strip.text = element_text(size = 16)
  )

theme_set(my_theme)

ggplot(plants_long) +
  geom_line(
    aes(time, height, col = treatment, group = interaction(plantid, treatment)),
    size = 2
    ) +
  facet_wrap(~ treatment)


plants_long |>
  filter(plantid == "plant_1")

```
