---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r}
library(tidyverse)
library(tidygraph)
library(ggraph)
```

1. [Movie Genres] How do movie genres relate to one another? Is romance + comedy
a more common combination than fantasy + adventure? We will use the dataset
[here](https://github.com/krisrs1128/stat436_s24/raw/main/data/movie_links.csv)
to answer these questions, using a node-link diagram inspired by (but much
simpler than) the [film flowers](https://observablehq.com/@sxywu/introduction-to-svg-and-d3-js) project
by Shirley Wu.

a. Build a `tbl_graph` containing the movies and their links.

```{r}
edges <- read_csv("https://github.com/krisrs1128/stat436_s24/raw/main/data/movie_links.csv")
genres <- read_csv("https://github.com/krisrs1128/stat436_s24/raw/main/data/genres.csv")

G <- tbl_graph(edges = edges) |>
    mutate(is_genre = name %in% genres$genre)
G
```

b. Create a node-link diagram to represent the connections between movies and
their genres.

```{r}
ggraph(G, "kk") +
    geom_edge_link(edge_color = "#d3d3d4") +
    geom_node_text(aes(label = name, size = is_genre)) +
    scale_size_discrete(range = c(3.5, 8)) +
    theme_void()

```


c. A list of all the genres in this network is available
[here](https://github.com/krisrs1128/stat992_f23/blob/main/activities/week9/genres.csv).
Design a strategy to distinguish between genres and movies, and justify your
choice of visual encodings. _Hint: Mutate the nodes in the `tbl_graph`._



1. [UK Energy Flow] We will visualize how energy is produced and consumed in the
United Kingdom, using a [dataset](https://github.com/krisrs1128/stat436_s24/raw/main/data/energy.csv) from the UK's Department of Energy and Climate
Change. Nodes and edges are used to classify types of activity that produce or
consume energy. For example, the edge

```
Wind    Electricity grid    289 
```

means that 289 KwH produced by wind power were sent to the country's electricity grid.

a. Build a `tbl_graph` from the provided edges. Ensure that the resulting
graph object has directed edges.

```{r}
edges <- read_csv("https://github.com/krisrs1128/stat436_s24/raw/main/data/energy.csv")
G <- tbl_graph(edges = edges)
G
```

b. Use `ggraph` to visualize the flow of energy across nodes in this
network. Ensure nodes are annotated.

```{r}
ggraph(G, "kk") +
    geom_edge_link(
        aes(edge_width = value),
        arrow = arrow(length = unit(1, "cm")),
        end_cap = circle(6, "mm")
    ) +
    scale_edge_width(range = c(.5, 10)) +
    geom_node_label(aes(label = name), size = 6)
edges
```

c. Experiment with several layouts. Which do you find most useful? Is there
an alternative layout you would prefer? Explain your reasoning.

```{r, fig.width = ..., fig.height = ...}
ggraph(G, "matrix") +
    geom_edge_tile() +
    geom_node_text(aes(label = name), size = 3, x = -2.5, nudge_y = 0.5, hjust = 0) +
  geom_node_text(aes(label = name), size = 3, angle = 90, y = 0.5, nudge_x = -0.5, hjust = 0) +
  scale_y_reverse(expand = c(0, 0, 0, 2.7)) + # make sure the labels aren't hidden
  scale_x_discrete(expand = c(0, 3, 0, 0)) +
    coord_fixed()
```

For labeling in adjacency matrices, see: https://github.com/krisrs1128/stat436_s24/blob/81d1f39b5e39088f8bbba32594f8b2bb73408153/notes/2024-12-27-week08-03.Rmd