---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(tsibble)
library(feasts)
theme_set(theme_bw())
```

[Global Carbon Footprint Network I] This problem studies a dataset of global
carbon emissions collected by the Global Carbon Footprint Network. It was the
subject of a previous [course
project](https://observablehq.com/@wang2274/milestone-4-team-4). We will first
quality check and structure the raw data. Then, we'll try answering some time
series questions visually.

a. The first few rows of the dataset are printed below. Write down 
two interesting questions that you think the dataset could help to answer.
Refer to at least one of the following vocabulary terms: trend, seasonality,
autocorrelation, cross-correlation, cyclic behavior.
    
```{r}
carbon <- read_csv("https://go.wisc.edu/7qx7u1")
head(carbon)
```
  
b. Build a `tsibble` object from the data above. Discuss why it can be
useful to convert the original `data.frame` object to this new `tsibble`
class.

```{r}
carbon <- carbon |>
  pivot_longer(-c("country", "Country Code"), names_to = "year")
```

This doesn't work, because year is a string.
```{r, eval = FALSE}
as_tsibble(carbon, index = year, key = country)
```

This still doesn't work -- there are duplicates!
```{r, eval = FALSE}
carbon |>
  mutate(year = as.integer(year)) |>
  as_tsibble(index = year, key = country)
```

For some reason, there are duplicated entries for Switzerland. This is one of
the reasons for using something like `tsibble`.
```{r}
carbon |>
  mutate(year = as.integer(year)) |>
  duplicates(key = country, index = year) |>
  arrange(year)
```

This code finally makes the tsibble.
```{r}
carbon_ts <- carbon |>
  filter(country != "Switzerland") |>
  mutate(year = as.integer(year)) |>
  as_tsibble(index = year, key = country)
```

c. Create either a line or heatmap visualization of the time series across all
countries. What have you learned from your visualization?

```{r}
ggplot(carbon_ts) +
  geom_line(aes(year, value, group = country))
```

```{r, fig.height = 12, fig.width = 6}
ggplot(carbon_ts) +
  geom_tile(aes(year, reorder(country, value), fill = log(1 + value), col = log(1 + value))) +
  scale_fill_distiller(direction = 1) +
  scale_color_distiller(direction = 1) +
  theme(axis.text.y = element_text(size = 6)) +
  scale_x_continuous(expand = c(0, 0))
```

I was surprised, but it seems that the Bahamas really did have a period of
intense carbon emissions due to [petroleum exploration](https://doi.org/10.1017/S0010417517000184).

d. Create a lag plot of the series for the United States. How do you
interpret this figure?

```{r}
carbon_ts |>
  filter(country == "United States") |>
  gg_lag()
```

This plots each timepoint against the time series' value $n$ steps in the past.
Each facet corresponds to a different $n$. Notice that there is stronger
correlation with the immediate past (lag 1) compared to the long-run past.

e. Create a lineplot across all countries after taking a moving aver
neighboring years. What are the advantages and disadvantages of smoothing
the original line plot?

This is a more complicated operation (I don't expect you to remember this), but
sometimes it comes in handy.

```{r}
library(slider)
smoothed_carbon <- carbon_ts |>
  group_by(country) |>
  mutate(smoothed = slide_dbl(value, ~ mean(.), .before = 5, .after = 5))
ggplot(smoothed_carbon) +
  geom_line(aes(year, smoothed, group = country))
```