---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r}
library(tidyverse)
library(tidygraph)
library(ggraph)
```

1. [Movie Genres] How do movie genres relate to one another? Is romance + comedy
a more common combination than fantasy + adventure? We will use the dataset
[here](https://github.com/krisrs1128/stat436_s24/raw/main/data/movie_links.csv)
to answer these questions, using a node-link diagram inspired by (but much
simpler than) the [film
flowers](https://observablehq.com/@sxywu/introduction-to-svg-and-d3-js) project
by Shirley Wu.

    a. Build a `tbl_graph` containing the movies and their links.
    
    b. Create a node-link diagram to represent the connections between movies and
    their genres.
    
    c. A list of all the genres in this network is available
    [here](https://github.com/krisrs1128/stat992_f23/blob/main/activities/week9/genres.csv).
    Design a strategy to distinguish between genres and movies, and justify your
    choice of visual encodings. _Hint: Mutate the nodes in the `tbl_graph`._

```{r}
edges <- read_csv("https://github.com/krisrs1128/stat436_s24/raw/main/data/movie_links.csv")
genres <- read_csv("https://github.com/krisrs1128/stat436_s24/raw/main/data/genres.csv")

G <- tbl_graph(edges = edges) %N>%
  mutate(type = name %in% genres$genre)

ggraph(G, "kk") +
     geom_edge_link(colour = "#d3d3d3", width = 0.5, alpha = 0.6) +
     geom_node_text(aes(label = name, size = type, col = type)) +
     scale_color_manual(values = c("#BF4545", "#225C73")) +
     scale_size_discrete(range = c(2.5, 5)) +
     theme_void() +
     theme(legend.position = "none")

ggraph(G, "matrix") +
  geom_edge_tile(mirror = TRUE) +
  geom_node_text(aes(label = name, col = type), x = -.5, hjust = 1) +
  geom_node_text(aes(label = name, col = type), y = -0.5, angle = 90, hjust = 1) +
  scale_x_discrete(expand = c(0, 200, 0, 0)) +
  scale_y_discrete(expand = c(0, 200, 0, 0)) +
  coord_fixed()
```

1. [UK Energy Flow] We will visualize how energy is produced and consumed in the
United States, using a [dataset](https://github.com/krisrs1128/stat436_s24/raw/main/data/energy.csv) from the UK's Department of Energy and Climate
Change. Nodes and edges are used to classify types of activity that produce or
consume energy. For example, the edge

```
Wind    Electricity grid    289 
```

means that 289 KwH produced by wind power were sent to the country's electricity grid.

a. Build a `tbl_graph` from the provided edges. Ensure that the resulting
graph object has directed edges.

b. Use `ggraph` to visualize the flow of energy across nodes in this
network. Ensure nodes are annotated.

c. Experiment with several layouts. Which do you find most useful? Is there
an alternative layout you would prefer? Explain your reasoning.



```{r}
edges <- read_csv("https://github.com/krisrs1128/stat436_s24/raw/main/data/energy.csv")
G <- as_tbl_graph(edges)
```

```{r, fig.height = 10, fig.width = 10}
ggraph(G, "kk") +
  geom_edge_link(
    aes(width = value), 
    arrow = arrow(length = unit(4, "mm")), 
    end_cap = circle(4, "mm")
  ) + 
  geom_node_label(aes(label = name))
```
