---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```
```{r}
library(tidyverse)
library(caret)
library(DALEX)
library(tidyverse)

th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Gender Pay Gap] This problem uses CP profiles to investigate the gender gap in
a Glassdoor dataset of employee salaries. It is helpful to use a model, because
it allows us to control for multiple other factors -- a direct plot of salary
vs. gender could be criticized as not accounting for confounding variables. The
code below trains a gradient boosting machine model on `BasePay` variable
(yearly salary in USD), using all potential predictors in the dataset.
    
```{r, warning = FALSE, message = FALSE}
salary <- read_csv("https://github.com/krisrs1128/stat479_s22/raw/main/_slides/week12/exercises/Glassdoor%20Gender%20Pay%20Gap.csv")

x <- salary |>
  select(Gender:Seniority) |>
  mutate(across(where(is.character), as.factor)) # gbm needs chr -> factor
y <- salary$BasePay
fit <- train(x, y, method = "gbm", verbose = FALSE)
```

a. Before attempting to explain the model, it is helpful to consider its
accuracy. Make a plot of the truth (`y`) against model predictions (`y_hat <- predict(fit)`) and comment on model performance.

b. Compute aggregate CP profiles grouped by gender and comment on the extent
of the gender pay gap. According to the fitted prediction surface, is there
more or less of a pay gap at certain ages or levels of seniority?

c. Show the analogous display without aggregating (i.e., `geom = "profiles"`). What is the interpretation of each line in this plot?
