---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

[Interactive German Traffic] This problem will revisit the previous problem
from an interactive point of view. We will build a visualization that helps
users explore daily traffic patterns across multiple German cities, using
interactivity to help users navigate the collection. We will need additional
features related to the day of the week for each timepoint, created by the
`wday` function below,

```{r}
library(tidyverse)
library(lubridate)
library(shiny)
theme_set(theme_bw())

plot_traffic <- function(df) {
  ggplot(df) +
    geom_line(aes(date, value, group = name)) +
    labs(x = "Date", y = "Traffic") +
    theme(
      #panel.background = element_rect(fill= "transparent"),
      #plot.background = element_rect(fill = "transparent", color = NA),
      axis.title = element_text(size = 20)
    )
}

traffic <- read_csv("https://uwmadison.box.com/shared/static/x0mp3rhhic78vufsxtgrwencchmghbdf.csv") |>
 mutate(day_of_week = wday(date))
```

a. Design and implement a Shiny app that allows users to visualize traffic
over time across selected subsets of cities. Make sure that it is possible
to view data from more than one city at a time. It is not necessary to label
the cities within the associated figure.

b. Introduce new inputs to allow users to select a contiguous range of days
of the week. For example, the user should have a way of zooming into the
samples taken within the Monday - Wednesday range.

First, let's experiment with a different layout. We'll also add some text
describing where the data come from and how to use the app.

```{r}
descriptive_text <- p("Explore traffic patterns over six days in 2016, across 38 German cities. These data were downloaded from a", a("Ridgeline Plot demo", href="https://observablehq.com/@d3/ridgeline-plot"), "by Mike Bostock, using data originally shared by Moritz Klick. Select a city in the dropdown menu to begin. Drag the slider endpoints to focus on a time range of interest")
```

```{r, eval = FALSE}
ui <- fluidPage(
  titlePanel("City Traffic Explorer"),
  selectInput("city", "City", unique(traffic$name), multiple = TRUE),
  sliderInput("day_of_week", "Days", 2, 7, c(2, 7)),
  plotOutput("time_series")
)

server <- function(input, output) {
  output$time_series <- renderPlot({
    traffic |>
      filter(
        name %in% input$city, 
        day_of_week >= input$day_of_week[1] & day_of_week <= input$day_of_week[2]
      ) |>
      plot_traffic()
  })
}

shinyApp(ui, server)
```

Next, let's experiment with some custom themes.

```{r, eval = FALSE}
library(bslib)

ui <- fluidPage(
  titlePanel("City Traffic Explorer"),
  sidebarLayout(
    sidebarPanel(
      descriptive_text,
      selectInput("city", "City", unique(traffic$name), multiple = TRUE),
      sliderInput("day_of_week", "Days", 2, 7, c(2, 7)),
    ),
    mainPanel(plotOutput("time_series"))
  )
)

server <- function(input, output) {
  output$time_series <- renderPlot({
    traffic |>
      filter(
        name %in% input$city, 
        day_of_week >= input$day_of_week[1] & day_of_week <= input$day_of_week[2]
      ) |>
      plot_traffic()
  }) # use bg = "transparent" to make the plot transparent
}

shinyApp(ui, server)
```