---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(embed)
library(tidymodels)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Single-Cell Genomics]  In this problem, we will apply UMAP to a
[dataset](https://uwmadison.box.com/shared/static/ai539s30rjsw5ke4vxbjrxjaiihq7edk.csv)
of Peripheral Blood Mononuclear Cells (PBMC) released by 10X Genomics. The first
column, `cell_tag`, gives an identifier for each cell in the dataset. All other
columns are molecules that were detected in that cell. For example, CD74 is a
molecule often found on the surface of T-cells.
    
a. Define a tidymodels `recipe` that specifies that UMAP should be performed
with the parameters `learn_rate = 0.1` and `neighbors = 5`. There is no need to
normalize these data, as they have been normalized in advance using methods
tailored to single-cell genomics data.

```{r}
pbmc <- read_csv("https://uwmadison.box.com/shared/static/ai539s30rjsw5ke4vxbjrxjaiihq7edk.csv")
# umap_rec <- 
```

b. Compute the UMAP embeddings across cells. Color points in by their value of
the GNLY molecule.

