---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(tsibble)
library(feasts)
library(plotly)
library(slider)
library(shiny)
theme_set(theme_bw())
```

[Global Carbon Footprint Network II] This problem creates interactive versions
of the carbon emissions dataset introduced in the previous problem.

a. Using `ggplotly`, create a line plot where country names can be revelaed
on hover.

This code is copied from the previous demo.

```{r}
carbon <- read_csv("https://go.wisc.edu/7qx7u1")
carbon_ts <- carbon |>
  pivot_longer(-c("country", "Country Code"), names_to = "year") |>
  filter(country != "Switzerland") |>
  mutate(year = as.integer(year)) |>
  as_tsibble(index = year, key = country)

smoothed_carbon <- carbon_ts |>
  group_by(country) |>
  mutate(smoothed = slide_dbl(value, ~ mean(.), .before = 5, .after = 5)) |>
  ungroup()

smoothed_carbon
```

b. Propose an interactive time series visualization that graphically queries a
time series feature (or several features) that can be generated using the
`feasts` package. What questions is your visualization particularly well-suited
to? What questions is it not appropriate for?

```{r}
trends <- smoothed_carbon |> 
  features(value, features = feature_set(tags = "trend"))

carbon_trends <- smoothed_carbon |>
  left_join(trends)
```

c. Implement your proposal from part (b).

```{r}
ui <- fluidPage(
  fluidRow(
    column(6, plotOutput("scatter", brush = "brush")),
    column(6, plotlyOutput("trends"))
  )
)

server <- function(input, output) {
  selected <- reactiveVal(rep(FALSE, nrow(carbon_trends)))
  observeEvent(
    input$brush, {
      brushed_result <- brushedPoints(carbon_trends, input$brush, allRows = TRUE)
      selected(brushed_result$selected_)
  })

  mutated_trends <- reactive({
    carbon_trends |>
      mutate(selected = selected())
  })
  
  output$scatter <- renderPlot({
    ggplot(mutated_trends()) +
      geom_point(aes(trend_strength, curvature, col = factor(selected)))
  })
  
  output$trends <- renderPlotly({
    current_data <- mutated_trends() |>
      filter(selected)

    if (nrow(current_data) > 0) {
      p <- mutated_trends() |>
        filter(selected) |>
        ggplot() +
        geom_line(aes(year, smoothed, group = country))
      return(ggplotly(p))
    } else {
      p <- mutated_trends() |>
        ggplot() +
        geom_line(aes(year, smoothed, group = country))
      return(ggplotly(p))
    }
  })
}

shinyApp(ui, server)
```