---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

[NCAA Trends] This [538
article](https://fivethirtyeight.com/features/louisiana-tech-was-the-uconn-of-the-80s/)
describes NCAA women's college basketball team performance over time. Each team
was assigned a score representing how successfully it played during each year
from 1982 to the present. This overall score is contained in the "points" column
below.

```{r}
ncaa <- read_csv("https://github.com/krisrs1128/stat992_f23/raw/main/exercises/ps1/ncaa_filtered.csv") 
```

a. Derive new columns representing (i) the cumulative total number of points
over time for each school (ii) the cumulative total number of points over
time for a hypothetical team that earns 35 points a year.

b. Create a visualization that shows the running total number of points for
each school over time. If you use faceting, ensure that facets are sorted in
an informative way.

c. Design a visualization that that compares each school's performance with
that of the hypothetical team that averages 35 points per year. See the
figure below for an example approach. Explain the strengths and weaknesses
of your design and comment on a finding from your visualization.

d. Note that the original data includes 250+ schools. Propose, but do not
implement, a visualization of the full dataset that makes use of dynamic
queries. What questions would the visualization answer? What would be the
structure of interaction, and how would the display update when the user
provides a cue?
