---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(shiny)
```

[Improving an app] Make the following apps more concise, modular, and
readable by using reactive expressions to capture duplicated computation and /
or externalizing complex computations into functions which are defined outside
of the server.

a. Program (a)

```{r, eval = FALSE}
ui <- fluidPage(
  titlePanel("Calculator"),
  numericInput("x", "Enter the value of x", 0),
  textOutput("f1"),
  textOutput("f2"),
  textOutput("f3")
)

server <- function(input, output) {
  output$f1 <- renderText({ 3 * input$x ^ 2 + 3})
  output$f2 <- renderText({ sqrt(3 * input$x ^ 2 + 3) - 5})
  output$f3 <- renderText({ 30 * input$x ^ 2 + 30})
}
    
shinyApp(ui, server)
```

Our revised version of this app is below. We use the fact that the three outputs
use the same intermediate calculation.

```{r}
ui <- fluidPage(
  titlePanel("Calculator"),
  numericInput("x", "Enter the value of x", 0),
  textOutput("f1"),
  textOutput("f2"),
  textOutput("f3")
)

server <- function(input, output) {
  f <- reactive({
    3 * input$x ^ 2 + 3
  })
  
  output$f1 <- renderText(f())
  output$f2 <- renderText({ sqrt(f()) - 5})
  output$f3 <- renderText(10 * f())
}
    
shinyApp(ui, server)
```


b. Program (b)
    
```{r, eval = FALSE}
penguins <- read_csv("https://uwmadison.box.com/shared/static/ijh7iipc9ect1jf0z8qa2n3j7dgem1gh.csv")
islands <- unique(penguins$island)
species <- unique(penguins$species)

ui <- fluidPage(
  titlePanel("Penguins Plot"),
  selectInput("species", "Species", species, multiple = TRUE),
  selectInput("island", "Island", islands, multiple = TRUE),
  selectInput("var1", "First Variable", colnames(penguins)[3:6]),
  selectInput("var2", "Second Variable", colnames(penguins)[3:6]),
  plotOutput("scatterplot"),
  plotOutput("histogram1"),
  plotOutput("histogram2"),
)

server <- function(input, output) {
  output$scatterplot <- renderPlot({
    current_data <- penguins |>
      filter(
        island %in% input$island,
        species %in% input$species
      )
    ggplot(current_data) +
      geom_point(aes(.data[[input$var1]], .data[[input$var2]]))
  })
  
  output$histogram1 <- renderPlot({
    current_data <- penguins |>
      filter(
        island %in% input$island,
        species %in% input$species
      )
    ggplot(current_data) +
      geom_histogram(aes(.data[[input$var1]]))
  })
  
  output$histogram2 <- renderPlot({
    current_data <- penguins |>
      filter(
        island %in% input$island,
        species %in% input$species
      )
    ggplot(current_data) +
      geom_histogram(aes(.data[[input$var2]]))
  })
  
}

shinyApp(ui, server)
```

The data are being filtered in almost the exact the same way within each
`renderPlot`. Therefore, it makes sense to introduce a new reactive element. We
used a trick to use string variables to specify columns within `aes`.
Specifically, we can refer to the original column using `.data[[v]]`, where `v`
is a variable storing the column name we want to encode.

```{r}
histogram <- function(df, v) {
  ggplot(df) +
    geom_histogram(aes(.data[[v]]))
}

ui <- fluidPage(
  titlePanel("Penguins Plot"),
  selectInput("species", "Species", species, multiple = TRUE),
  selectInput("island", "Island", islands, multiple = TRUE),
  selectInput("var1", "First Variable", colnames(penguins)[3:6]),
  selectInput("var2", "Second Variable", colnames(penguins)[3:6]),
  plotOutput("scatterplot"),
  plotOutput("histogram1"),
  plotOutput("histogram2"),
)

server <- function(input, output) {
  current_data <- reactive({
    penguins |>
      filter(
        island %in% input$island,
        species %in% input$species
      )
  })
  
  output$scatterplot <- renderPlot({
    ggplot(current_data()) +
      geom_point(aes(.data[[input$var1]], .data[[input$var2]]))
  })
  
  output$histogram1 <- renderPlot(histogram(current_data(), input$var1))
  output$histogram2 <- renderPlot(histogram(current_data(), input$var2))
}

shinyApp(ui, server)
```
