---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(shiny)
library(sf)
library(leaflet)
theme_set(theme_bw())
```

[Glacial Lakes]. The data at this
[link](https://raw.githubusercontent.com/krisrs1128/stat436_s23/main/data/GL_3basins_2015.topojson)
contain labels of glacial lakes the Hindu Kush Himalaya, created during an
ecological survey in 2015 by the International Centre for Integrated Mountain
Development.

```{r}
lakes <- read_sf("https://raw.githubusercontent.com/krisrs1128/stat436_s23/main/data/GL_3basins_2015.topojson") |>
  mutate(log_area = log(Area))
cols <- c("#023859", "#99ADBF")

```

d. Part (c) allowed us to visualize the lake data within a very small region.
How might you use interactivity to navigate the full collection of lakes?
Describe at least one query that allows us to filter by a property of the lake
and another that allows us to filter by spatial extent. How would you lay out
the visualization components to streamline these queries?

e. Implement an interactive visualization with one of the queries discussed in
part (d).

```{r}
ui <- fluidPage(
  fluidRow(
    column(6, plotOutput("histogram", brush = brushOpts("plot_brush", direction = "x"))),
    column(6, leafletOutput("map"))
  ),
  dataTableOutput("table")
)

server <- function(input, output) {
  # setup the brush interaction
  selected <- reactiveVal(rep(TRUE, nrow(lakes)))
  observeEvent(
    input$plot_brush, {
      selected(brushedPoints(lakes, input$plot_brush, allRows = TRUE)$selected_)
    }
  )
  
  # draw the histogram
  output$histogram <- renderPlot({
    ggplot(lakes, aes(log_area)) +
      geom_histogram(fill = cols[2]) +
      geom_histogram(data = filter(lakes, selected()), fill = cols[1]) +
      scale_y_continuous(expand = c(0, 0, 0, 0.9))
  })
  
  # filter the table based on interactions
  output$table <- renderDataTable({
    current_lakes <- lakes |>
      filter(selected())
    
    current_lakes |>
      arrange(-Area)
  })
  
  # initial map render
  output$map <- renderLeaflet({
    ###
  })
}

shinyApp(ui, server)
```
