---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)

th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Histopathology Embeddings] This problem investigates the features learned by
a residual network model trained trained to classify histopathology slides.
Specifically, the script at [this
link](https://colab.research.google.com/drive/1bPFz718F-YE0iUcdkm2kc0Aeg8g6-Neh?usp=sharing)
was used to train a model to images from the Pcam [benchmark
dataset](https://www.kaggle.com/competitions/histopathologic-cancer-detection/data).
Each image is a tissue slide. The class labels describe whether the center $32 \times 32$ patch within the image contains any cancerous cells.

In the process, we will also practice using the `reticulate` package to read in
numpy arrays produced by the python training script linked above. This language
interoperability makes it possible to use the packages best suited to both
modeling (`pytorch`) and visualization (`ggplot2`).

a. We have hosted a subsample of the training images at [this
link](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week13/exercises/subset.tar.gz?raw=true).
Their corresponding [labels](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week13/exercises/y.npy?raw=true) and [file names](https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week13/exercises/fnames.npy) are stored as numpy arrays. Visualize the raw images corresponding to 10 images from each class. _Hint: To unzip these files from the command line, you can use `tar -zxvf subset.tar.gz`_  

```{python}
import numpy as np
fnames = np.load("fnames.npy")
y = np.load("y.npy")
```

```{r}
library(fs)
library(raster)
library(reticulate)
library(tmap)

# download the subset.tar.gz and unzip
data_dir <- path("/Users/ksankaran/Downloads/subset/")
ix <- which(py$y == 1)[1:10]
ims <- map(py$fnames[ix], ~ brick(data_dir / .))
map(ims, ~ tm_shape(.) + tm_rgb())
```

```{r}
ix <- which(py$y == 0)[1:10]
ims <- map(py$fnames[ix], ~ brick(data_dir / .))
map(ims, ~ tm_shape(.) + tm_rgb())
```

b. For the subsample in part (a), we have saved the residual network
features from the final pre-classification layer. They are available
at [this
link](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week13/exercises/h.npy?raw=true).
Generate UMAP embeddings for the images based on these features, and shade
in each sample according to its class.

```{python}
h = np.load("h.npy")
```

```{r}
library(tidymodels)
library(embed)
umap_rec <- data.frame(py$h) |>
  mutate(
    file = py$fnames,
    y = as.factor(py$y)
  ) |>
  recipe(~ ., data = .) |>
  update_role(y, file, new_role = "id") |>
  step_normalize(all_predictors()) |>
  step_umap(all_predictors())

umap_prep <- prep(umap_rec)
coords <- juice(umap_prep)
```

```{r, fig.width = 10, fig.height = 10}
p <- ggplot(coords) +
  geom_point(aes(UMAP1, UMAP2, col = y)) +
  coord_fixed()
p
```

c. Using `annotation_raster` layers from `ggplot2`, display the original
images from (a) at the locations of the UMAP coordinates from (b). The
correspondence between image filenames and features is given by [this
array](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week13/exercises/fnames.npy?raw=true).
In particular, the $i^{th}$ element of this array is the source image for
the $i^{th}$ row of the features matrix.

```{r, fig.width = 10, fig.height = 10}
library(progress)
pb <- progress_bar$new()

s <- .1
for (i in seq_len(nrow(coords))) {
  im <- brick(data_dir / path(coords$file[i]))
  p <- p + 
    annotation_raster(
      as.array(im) / 255, 
      xmin = coords$UMAP1[i] - s, 
      xmax = coords$UMAP1[i] + s, 
      ymin = coords$UMAP2[i] - s, 
      ymax = coords$UMAP2[i] + s
    )
  
  pb$tick()
}

p
#ggsave("~/Downloads/image.png", width = 12, height = 12, dpi = 500)
```
