---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(topicmodels)
library(tidytext)

th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Hotel Reviews] In this problem, we will practice using Latent Dirichlet
Allocation to understand the topics that come up across hotel reviews from an
online database. We will also study whether there are certain topics that are
more common in positive vs. negative reviews.

a. Transform the raw reviews into a collection of per-review word counts.
Remove stopwords from across all lexicons in `tidytext::stopwords`.

```{r}
reviews <- read_csv("https://github.com/krisrs1128/stat436_s24/raw/main/data/reviews.csv")
reviews
reviews_dtm <- reviews |>
  unnest_tokens(word, Review) |>
  filter(!(word %in% stop_words$word)) |>
  count(document, word) |>
  cast_dtm(document, word, n)
```
    
b. Convert the `data.frame` from (a) into a topic models Document-Term
matrix (i.e., an object of class `DocumentTermMatrix`). Fit an LDA model
with 8 topics to the prepared object.

```{r}
fit <- LDA(reviews_dtm, k = 8)
```

c. Create a heatmap of topic distributions for the 50 words $d$ with the
largest inter-quartile across topics (i.e., largest IQR for
$\left(\beta_{dk}\right)_{k = 1}^{8}\right)$. Provide an interpretation for
one of the topics, based on the words that have high probability within it.

```{r}
## get the topic matrix
beta <- tidy(fit, matrix = "beta")
head(beta)

## pivot the topic matrix wider
beta_wide <- beta |>
  pivot_wider(names_from = "topic", values_from = "beta") |>
  column_to_rownames("term")

term_ordering <- beta |>
  group_by(term) |>
  summarise(
    range = IQR(beta)
  ) |>
  arrange(-range) |>
  pull(term)

## use superheat library
#install.packages("superheat")
library(superheat)
superheat(
  beta_wide[term_ordering[1:40], ],
  pretty.order.rows = TRUE,
  pretty.order.cols =  TRUE
)
```

d. Create a Structure plot displaying the topic memberships for each review.
Sort reviews according to their order on a hierarchical clustering tree, and
facet documents according to hotel rating. Are there certain topics that
appear to be more common in negative vs. positive reviews?

```{r}
# sort the documents according to their membership
hclust_result <- hclust(dist(fit@gamma))
document_order <- fit@documents[hclust_result$order]

# get the memberships
memberships <- tidy(fit, matrix = "gamma") |>
  left_join(
    reviews |>
      mutate(document = as.character(document))
    ) |>
  mutate(document = factor(document, levels = document_order))



# make stacked area/bar plot
ggplot(memberships) +
  geom_col(aes(document, gamma, fill = factor(topic))) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    axis.ticks = element_blank(),
    axis.text.x = element_blank()
  ) +
  scale_fill_brewer(palette = "Set2") +
  facet_grid(. ~ Rating, scales = "free_x", space = "free_x")



```