---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```


[Plant Growth Experiment] This problem will give you practice with tidying a
dataset so that it can be easily visualized. The data describe the height of
several plants measured every 7 days. The plants have been treated with
different amounts of a growth stimulant. The first few rows are printed below --
`height.x` denotes the height of the plant on day `x`.

```{r}
library(tidyverse)
plants <- read_csv("https://uwmadison.box.com/shared/static/qg9gwk2ldjdtcmmmiropcunf34ddonya.csv")
plants
```

a. Propose an alternative arrangement of rows and columns that conforms to the
tidy data principle.

According to the tidy data principle, each sample should have its own row. Right
now, however, there are four samples on each row -- one for each week of the
experiment. We need to reshape the data so that each of these four measurements
appears on its own row.

b. Implement your proposed arrangement from part (a).

```{r}
plants_reshaped <- plants %>%
  pivot_longer(starts_with("height")) %>%
  mutate(day = str_extract(name, "[0-9]+"))
```


c. Using the dataset from (b), design and implement a visualization showing the
growth of the plants over time according to different treatments.

```{r}
ggplot(plants_reshaped) +
  geom_line(aes(day, value, group = plantid)) +
  facet_wrap(~ treatment)
```

```{r}
ggplot(plants_reshaped) +
  geom_line(aes(day, value, group = interaction(plantid, treatment), col = treatment))
```
