---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(sf)
library(tmap)
library(tidyverse)
library(ceramic)
theme_set(theme_bw())
```

[Glacial Lakes]. The data at this [link](https://raw.githubusercontent.com/krisrs1128/stat436_s23/main/data/GL_3basins_2015.topojson) contain labels of glacial lakes the
Hindu Kush Himalaya, created during an ecological survey in 2015 by the
International Centre for Integrated Mountain Development.

```{r}
lakes <- read_sf("https://raw.githubusercontent.com/krisrs1128/stat436_s23/main/data/GL_3basins_2015.topojson") %>%
  mutate(log_area = log(Area))

cols <- c("#023859", "#99ADBF")
```

a. How many lakes are in this dataset? What are the latitude / longitude
coordinates of the largest lakes in each Sub-basin?

There are `r nrow(lakes)` lakes in this dataset.

```{r}
top_lakes <- lakes %>%
  group_by(Sub_Basin) %>%
  slice_max(Area)
```

b. Plot the polygons associated with each of the lakes identified in step
(a). _Hint: You may find it useful to split lakes across panels using the
`tmap_facets` function. If you use this approach, make sure to include a
scale with `tm_scale_bar()`, so that it is clear that lakes have different
sizes.

```{r}
tm_shape(top_lakes) +
  tm_polygons() +
  tm_facets("Sub_Basin") +
  tm_scale_bar()
```

Let's refine this figure. One improvement is to sort the lakes by decreasing
size. We can't use `reorder()` like in the usual ggplot `facet` commands, but we
can manually set the factors levels to sort the faceting panels.

```{r}
basin_order <- top_lakes %>%
  arrange(-Area) %>%
  pull(Sub_Basin) 

top_lakes <- top_lakes %>%
  mutate(Sub_Basin = factor(Sub_Basin, levels = basin_order))
```

The final refinement is to change the lake color and add a scale for reference.

```{r}
tm_shape(top_lakes) +
  tm_polygons(col = cols[1]) +
  tm_facets("Sub_Basin") +
  tm_scale_bar()
```

c. Visualize all lakes with latitude between 28.2 and 28.4 and with
longitude between 85.8 and 86. Optionally, add a basemap associated with
each lake.

```{r}
lakes_sub <- lakes %>%
  filter(
    Latitude > 28.2, Latitude < 28.4,
    Longitude > 85.8, Longitude < 86
  )

tm_shape(lakes_sub) +
  tm_polygons(col = cols[1])
```

```{r}
#Sys.setenv(MAPBOX_API_KEY="pk.eyJ1Ijoia3Jpc3JzMTEyOCIsImEiOiJjbDYzdjJzczQya3JzM2Jtb2E0NWU1a3B3In0.Mk4-pmKi_klg3EKfTw-JbQ")

basemap <- cc_location(cbind(85.9, 28.3), buffer = 1e4)
tm_shape(basemap) +
  tm_rgb() +
  tm_shape(lakes_sub) +
  tm_polygons(col = cols[1])
```
