---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(tidymodels)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```


[Food nutrients] This problem will use PCA to provide a low-dimensional view
of a 14-dimensional nutritional facts
[dataset](https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv).
The data were originally curated by the USDA and are regularly used in
[visualization studies](http://bl.ocks.org/syntagmatic/raw/3150059/).
    
```{r}
nutrients <- read_csv("https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv")
```
    
a. Define a tidymodels `recipe` that normalizes all nutrient features and The
code below defines a recipe that first normalizes the predictors and then
performs PCA.

```{r}
pca_recipe <- recipe(~., data = nutrients) |>
  update_role(name, id, starts_with("group"), new_role = "id") |>
  step_normalize(all_predictors()) |>
  step_pca(all_predictors())
pca_prep <- prep(pca_recipe)
```

specifies that PCA should be performed.

b. Visualize the top 6 principal components. What types of food do you expect to
have low or high values for PC1 or PC2?

The first 6 components are plotted below. Each component can be thought of as a
derived feature or contrast that explains a significant amount of the variation
across the foods. Since the first PC has a very positive weight on water and a
very negative weight on calories and fat, we suspect that it distinguishes
between fruit / vegetables and everything else. Since the second PC has very
positive values for carbohydrates and very negative values for water and fats,
we suspect that these are starchy foods, like bread or pasta.

```{r, fig.width = 10, fig.height = 8}
pca_result <- tidy(pca_prep, 2) |>
  mutate(terms = str_replace(terms, " \\(g\\)", ""))
ggplot(pca_result |> filter(component %in% str_c("PC", 1:6))) +
  geom_col(aes(x = value, y = terms)) +
  facet_wrap(~ component) +
  labs(x = "Component", y = "Features")
```

c. Compute the average value of PC2 within each category of the `group` column.
Give the names of the groups sorted by this average.

The `bake` function below extracts the PC scores for each sample. We then
compute the mean of PC2 within each food group using the `group_by` and
`summarise` pattern.

```{r}
pca_scores <- bake(pca_prep, nutrients)
group_order <- pca_scores |>
  group_by(group) |>
  summarise(mpc2 = mean(PC2)) |>
  arrange(mpc2) |>
  pull(group)
```

d. Visualize the scores of each food item with respect to the first two
principal components. Facet the visualization according to the `group` column,
and sort the facets according to the results of part (c). How does the result
compare with your guess from part (b)?

We mutate the `group` column so that the facets will be reordered according to
the results of (c). In general, the facets of a ggplot2 plot will always be
ordered by the factor levels of the defining variable. From here, the result is
a standard scatterplot, though we make the axes slightly bolder and customize
the color and axes labels.

```{r, fig.width = 12}
pca_scores |>
  mutate(group = factor(group, levels = group_order)) |>
  ggplot(aes(x = PC1, y = PC2)) +
  geom_vline(xintercept = 0, col = "#4a4a4a") +
  geom_hline(yintercept = 0, col = "#4a4a4a") +
  geom_point(size = 0.4, alpha = 0.6) +
  scale_x_continuous(breaks = seq(-8, 0, length.out = 3)) +
  scale_color_brewer(palette = "Set2") +
  facet_wrap(~ reorder(group, PC1), ncol = 9) +
  coord_fixed() +
  theme(strip.text = element_text(size = 8))
```