---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

[Multiple Weather Variables] In this problem, we will construct a compound
figure to simultaneously visualize several weather variables in one display. The
block below reads a dataset that includes temperature, precipitation, and wind
measurements for New York City and Seattle over the course of several years. It
then then averages each variable for each day (1 - 365) in the year.

```{r}
library(tidyverse)
theme_set(theme_bw())

weather <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/exercises/data/weather.csv") |>
  mutate(day_in_year = lubridate::yday(date)) |>
  group_by(location, day_in_year) |>
  summarise(across(precipitation:wind, mean))
```
  
a. Construct three base `ggplot2` plots based on the variables that are
provided. For example, you may construct a line plot (`geom_line`) of
average precipitation, a histogram (`geom_histogram`) of per-city
windspeeds, or a ribbon plot (`geom_ribbon`) of temperature ranges over
time. Make sure to display at least two weather variables from among the
four that are provided.

```{r}
p <- list()
p[["ribbon"]] <- ggplot(weather) +
  geom_ribbon(aes(day_in_year, ymin = temp_min, ymax = temp_max, fill = location), alpha = 0.8) +
  scale_x_continuous(expand = c(0, 0))

p[["precip"]] <- ggplot(weather) +
  geom_point(aes(day_in_year, precipitation, col = location, size = wind)) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0, 0.1, 0)) +
  scale_size_continuous(range = c(0.1, 4))

p[["range"]] <- ggplot(weather) +
  geom_histogram(
    aes(temp_max - temp_min, fill = location), 
    alpha = 0.8, position = "identity"
  ) +
  scale_y_continuous(expand = c(0, 0, 0.1, 0))

p
```

b. Design a compound figure based on the base plots in part (a). Why did you
lay out the subplots in the locations that you did? Ensure that consistent
visual encodings are used throughout, that legends are collected, and that
clear but unobtrusive annotation is provided.

```{r, fig.width = 10, fig.height = 5}
library(patchwork)
p[["range"]] + (p[["ribbon"]] / p[["precip"]])
```

A more refined version collects redundant legends and provides sub-plot labels.

```{r}
p[["range"]] <- p[["range"]] +
  labs(x = "Temperature Range", title = "a")
p[["ribbon"]] <- p[["ribbon"]] +
  labs(x = "Day of Year", y = "Temperature", title = "b")
p[["precip"]] <- p[["precip"]] + 
  labs(x = NULL, title = "c") +
  scale_color_discrete(guide = FALSE)
p[["range"]] + (p[["ribbon"]] / p[["precip"]]) +
  plot_layout(guides = "collect", widths = c(1, 3))
```

c. Discuss, but do not implement, an alternative compound figure for the
same data, with either different base plots or overall layout (or both). For
which visual comparisons are either designs best suited?

For sub-plots (b - c), we could imagine grouping days of the year to just
months. For (b), we could replace the ribbon with the average minimum and
maximum temperatures. We could also use a boxplot for the bottom panel. The
advantage of this approach is that it would help denoising the two figures. For
example, the higher precipitation in New York during the summer would likely
become more obvious. However, it would make it impossible to see the degree of
day-to-day variability in these panels.