---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(tidymodels)
th <- theme_minimal() + 
  theme
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Food nutrients] This problem will use PCA to provide a low-dimensional view
of a 14-dimensional nutritional facts
[dataset](https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv).
The data were originally curated by the USDA and are regularly used in
[visualization studies](http://bl.ocks.org/syntagmatic/raw/3150059/).
    
```{r}
nutrients <- read_csv("https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv")
nutrients
```
    
a. Define a tidymodels `recipe` that normalizes all nutrient features and run PCA.

```{r}
pca_recipe <- recipe(~ ., data = nutrients) |>
  update_role(name, id, starts_with("group"), new_role = "id") |>
  step_normalize(all_predictors()) |>
  step_pca(all_predictors())

pca_prep <- prep(pca_recipe)
```

b. Visualize the top 6 principal components. What types of food do you expect to
have low or high values for PC1 or PC2?

```{r, fig.width = 10, fig.height = 8}
tidy(pca_prep, 2) |>
  filter(component %in% paste0("PC", 1:6)) |>
  ggplot() +
    geom_col(aes(value, terms)) +
    facet_wrap(~ component) +
    theme(
      axis.text = element_text(size = 16),
      strip.text = element_text(size = 16),
      axis.title = element_text(size = 20)
      )

```

c. Compute the average value of PC2 within each category of the `group` column.
Give the names of the groups sorted by this average.

```{r}
```

d. Visualize the scores of each food item with respect to the first two
principal components. Facet the visualization according to the `group` column,
and sort the facets according to the results of part (c). How does the result
compare with your guess from part (b)?

```{r}
theme_set(theme_classic())
scores <- bake(pca_prep, nutrients)
ggplot(scores) +
  geom_hline(yintercept = 0, linewidth = 1) +
  geom_vline(xintercept = 0, linewidth = 1) +
  geom_point(aes(PC1, PC2, col = group_lumped)) +
  scale_color_brewer(palette = "Set3") +
  facet_wrap(~ reorder(group, PC2))
```

