---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(sf)
library(tmap)
library(tidyverse)
library(ceramic)
theme_set(theme_bw())
```

[Glacial Lakes]. The data at this [link](https://raw.githubusercontent.com/krisrs1128/stat436_s23/main/data/GL_3basins_2015.topojson) contain labels of glacial lakes the
Hindu Kush Himalaya, created during an ecological survey in 2015 by the
International Centre for Integrated Mountain Development.

```{r}
lakes <- read_sf("https://raw.githubusercontent.com/krisrs1128/stat436_s23/main/data/GL_3basins_2015.topojson") |>
  mutate(log_area = log(Area))

cols <- c("#023859", "#99ADBF")
```

a. How many lakes are in this dataset? What are the latitude / longitude
coordinates of the largest lakes in each Sub-basin?

b. Plot the polygons associated with each of the lakes identified in step
(a). _Hint: You may find it useful to split lakes across panels using the
`tmap_facets` function. If you use this approach, make sure to include a
scale with `tm_scale_bar()`, so that it is clear that lakes have different
sizes.

c. Visualize all lakes with latitude between 28.2 and 28.4 and with
longitude between 85.8 and 86. Optionally, add a basemap associated with
each lake.

```{r}
#Sys.setenv(MAPBOX_API_KEY="")
#basemap <- cc_location(cbind(85.9, 28.3), buffer = 1e4)
```
