---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Women, Business and the Law] The World Bank's Women, Business, and the Law
program has curated a dataset about gender equality across countries since 1971.
Since there are more than 30 variables summarizing each country for each year,
we will use clustering to observe the general trends across all measures.

a. The code below reads in the data, simplifies column names (saving the
original names in the `question_mapping` data.frame), and converts text Yes
/ No responses to numerical 0 / 1. Widen the dataset so that each row
corresponds to a single country $\times$ year combination.

```{r}
# read in and create a codebook
survey <- read_csv("https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week9/exercises/Viz5-July-2020-economic-empowerement.csv?raw=true")
question_mapping <- survey %>%
  select(Question, `Question Category`) %>%
  distinct() %>%
  mutate(q_id = str_c("Q", row_number()))

recode_vector <- setNames(question_mapping$Question, question_mapping$q_id)
survey <- survey %>%
  mutate(
    Question = fct_recode(Question, !!!recode_vector),
    answer = as.numeric(ifelse(`Text Answer` == "Yes", 1, 0))
  ) %>%
  select(Country, `WBL Report Year`, Question, answer)

survey <- survey %>%
  pivot_wider(names_from = "Question", values_from = "answer")
```

b. Apply $K$-means to the question responses (you may choose $K$). Visualize
the centroids and briefly interpret each cluster in context, using the
original text of the questions stored in `question_mapping`.

```{r}
# run kmeans
fit <- survey %>%
  select(starts_with("Q")) %>%
  kmeans(centers = 5)

# make it "long" again
centroids <- data.frame(fit$centers) %>%
  rownames_to_column("cluster") %>%
  pivot_longer(-cluster, names_to = "q_id")
```

```{r, fig.width = 10, fig.height = 8}
# make the faceted centroids means, sorting clusters according to averages
ggplot(centroids) +
  geom_col(aes(value, reorder(q_id, value, var)), width = 1) +
  facet_wrap(~ reorder(cluster, value), ncol = 5) +
  theme(axis.text.y = element_text(size = 10))
```
    
```{r, fig.height = 9, fig.width = 3}
ggplot(centroids) +
  geom_tile(aes(reorder(cluster, value), reorder(q_id, value), fill = value)) +
  scale_fill_gradient(low = "white", high = "black")
```

    
c. Visualize change in cluster sizes over time. Provide a brief
interpretation (for example, in which types of questions is there the most /
least progress towards equality?).

```{r, fig.width = 10, fig.height = 4}
interpret <- bind_cols(
  survey, 
  cluster = as.factor(fit$cluster)
) %>%
  count(`WBL Report Year`, cluster)

slope_fun <- function(x) {
  mean(diff(x))
}
  
ggplot(interpret) +
  geom_point(aes(`WBL Report Year`, n)) +
  facet_wrap(~ reorder(cluster, n, slope_fun))
```
