---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(tsibble)
library(feasts)
library(plotly)
library(slider)
library(shiny)
theme_set(theme_bw())
```

[Global Carbon Footprint Network II] This problem creates interactive versions
of the carbon emissions dataset introduced in the previous problem.

a. Using `ggplotly`, create a line plot where country names can be revelaed
on hover.

This code is copied from the previous demo.

```{r}
carbon <- read_csv("https://go.wisc.edu/7qx7u1")
carbon_ts <- carbon |>
  pivot_longer(-c("country", "Country Code"), names_to = "year") |>
  filter(country != "Switzerland") |>
  mutate(year = as.integer(year)) |>
  as_tsibble(index = year, key = country)

smoothed_carbon <- carbon_ts |>
  group_by(country) |>
  mutate(smoothed = slide_dbl(value, ~ mean(.), .before = 5, .after = 5)) |>
  ungroup()
```

We now construct a plotly object from the ggplot time series.

```{r}
p <- ggplot(smoothed_carbon) +
  geom_line(aes(year, smoothed, group = country))
ggplotly(p)
```

What if we want to be able to hover and get the country names?

b. Propose an interactive time series visualization that graphically queries a
time series feature (or several features) that can be generated using the
`feasts` package. What questions is your visualization particularly well-suited
to? What questions is it not appropriate for?

We will try querying series according to their `spikiness` and `curvature`. The
graphical query will allow us to highlight countries using a scatterplot of
these feature values.

c. Implement your proposal from part (b).

```{r}
trends <- smoothed_carbon |>
  features(value, features = feature_set(tags = "trend")) |>
  mutate(log_spikiness = log(spikiness))

carbon_trends <- smoothed_carbon |>
  left_join(trends)
```

```{r}
ui <- fluidPage(
  fluidRow(
    column(6, plotOutput("scatter", brush = "brush")),
    column(6, plotlyOutput("trends"))
  )
)

server <- function(input, output) {
  selected <- reactiveVal(rep(FALSE, nrow(carbon_trends)))
  observeEvent(
    input$brush, {
      selected(brushedPoints(carbon_trends, input$brush, allRows = TRUE)$selected_)
  })
  
  mutated_trends <- reactive({
    carbon_trends |>
      mutate(selected = selected())
  })
  
  output$scatter <- renderPlot({
    ggplot(mutated_trends()) +
      geom_point(aes(log_spikiness, curvature, col = selected), size = 2) +
      scale_color_manual(values = c("#B0D1D9", "#273859"), guide = "none")
  })
  
  output$trends <- renderPlotly({
    current <- mutated_trends()
    if (any(current$selected)) {
      p <- current |>
        filter(selected) |>
        ggplot() +
        geom_line(aes(year, smoothed, group = country), col = "#273859") +
        theme(legend.position = "none")
      ggplotly(p)
    }
  })
}

shinyApp(ui, server)
```