---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo = FALSE}
knitr::opts_chunk$set(warnings = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Taxi Trips] In this problem, we will use hierarchical clustering to find
typical taxi trip trajectories in Porto, Portugal. The data are a subset from
the the [ECML / PKDD 2015
Challenge](http://www.geolink.pt/ecmlpkdd2015-challenge/dataset.html) -- the
link provides a complete data dictionary. We have preprocessed it into two
formats. The first
([`wide`](https://uwmadison.box.com/shared/static/cv0lij4d9gn3s8m2k98t2ue34oz6sbj5.csv))
includes each taxi trip on its own row, with latitude and longitude coordinates
along the journey given as separate columns (`x_0, y_0` is the origin of the
trip and `x_15`, `y_15` is the destination). The second
([`long`](https://uwmadison.box.com/shared/static/098cjaetm8vy0mufq21mc8i9nue2rr2b.csv))
format spreads each point of the journey into a separate row.
    
a. Filter the `long` form of the data down to trip `1389986517620000047` and
plot its trajectory as a sequence of points.

```{r}
trips <- read_csv("https://uwmadison.box.com/shared/static/098cjaetm8vy0mufq21mc8i9nue2rr2b.csv", col_types = cols(TRIP_ID = col_character()))
trips_wide <- read_csv("https://uwmadison.box.com/shared/static/cv0lij4d9gn3s8m2k98t2ue34oz6sbj5.csv", col_types = cols(TRIP_ID = col_character()))

# test code on a subset
# trips <- trips |>
#   filter(row_number() < 1000)
# 
# trips_wide <- trips_wide |>
#   filter(TRIP_ID %in% trips$TRIP_ID)
```

The code below filters to the required trip and uses a scatterplot to show the
trajectory.

```{r, fig.width = 5, fig.height = 2}
trips |>
  filter(TRIP_ID == "1389986517620000047") |>
  ggplot(aes(x, y)) +
  geom_point() +
  coord_fixed()
```

    
b. We could hierarchically cluster rows in either the `wide` or the `long`
format datasets. How would the interpretation of the results differ between the
two approaches?

If we hierarchically clustered the `long` dataset, then a trip might be split
into different clusters. The resulting clusters would be related to the
geographical positions of each trip timepoint, rather than the shape of the
overall trajectory.

On the other hand, clustering the `wide` dataset will identify trajectories that
have similar overall shapes. For example, we would expect this approach to
identify main highways in the city, or common origin-destination pairs.

c. Compute a hierarchical clustering of the `wide` format data, using only the
columns starting with `x` or `y` as features.

The `dist` function below computes the pairwise distances between trajectories,
and this is then fed into a hierarchical clustering routine.

```{r}
hclust_res <- trips_wide |>
  select(-TRIP_ID) |>
  dist() |>
  hclust()
```

d. Cut the hierarchical clustering tree so that 8 clusters are produced.
Visualize the trajectories of the taxi trips either colored or faceted by their
cluster.

We add a `cluster` column to the wide trips dataset based on the result of (c).
By then joining this with the original long-format dataset, we can plot each
taxi trip coordinate with the associated trajectory cluster.

```{r, fig.cap = "An example result for problem 3(c). A small amount of jitter has been added, to make the larger highways more clearly visible.", fig.width = 8, dev = "png", dpi = 1200}
trips_wide <- trips_wide |>
  mutate(cluster = cutree(hclust_res, 8))

trips <- trips |>
  left_join(trips_wide)

ggplot(trips) +
  geom_point(
    aes(x = x, y = y, col = as.factor(cluster)),
    size = 0.1, alpha = 0.05, position = position_jitter(w = 0.0005, h = 0.0005)
  ) +
  scale_color_brewer(palette = "Set2") +
  coord_fixed() +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1))) +
  labs(x = "Longitude", y = "Latitude", col = "Cluster")
```
