---
title: "Patchwork"
description: |
  Implementing compound figures in R
author:
  - name: Kris Sankaran
layout: post
output:
  distill::distill_article:
    self_contained: false
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE, echo = TRUE, fig.height = 5, retina = 4)
```


_[Reading](https://clauswilke.com/dataviz/multi-panel-figures.html), [Recording](https://mediaspace.wisc.edu/media/Week%203%20%5B4%5D%20Patchwork/1_59ftmcdh),  [Rmarkdown](https://github.com/krisrs1128/stat479_s22/blob/main/_posts/2022-02-06-week03-04/week03-04.Rmd)_

1. In the last set of notes we discussed principles for designing effective
compound figures. In these notes, we'll review the `patchwork` R package, which
can be used to implement compound figures.

1. This package creates a simple syntax for combining figures,

    * `p1 + p2` concatenates two figures horizontally
    * `p1 / p2` concatenates two figures vertically
    
    This idea is simple, but becomes very powerful once we realize that we can
    define a whole algebra on plot layouts,
    
    * `p1 + p2 + p3` concatenates three figures horizontally
    * `p1 / p2 / p3` concatenates three figures vertically
    * `(p1 + p2) / p3` Concatenates the first two figures horizontally, and places the third below both.
    * ...

1. Before we illustrate the use of this package, let's read in the athletes data
from the previous notes. The code below constructs the three component plots
that we want to combine. Though it looks like a lot of code, it's just because
we are making several plots and styling each one of them. Conceptually, this is
the same type of `ggplot2` code that we have been using all semester -- the only
difference is that we save all the figure objects into one list, instead of
printing them right away.
    
    ```{r}
    library(tidyverse)
    library(patchwork)
    
    athletes <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat436_s23/main/data/athletes.csv") %>%
      filter(sport %in% c("basketball", "field", "rowing", "swimming", "tennis", "track (400m)")) %>%
      mutate(sex = recode(sex, "m" = "male", "f" = "female"))
    
    p <- list()
    p[["bar"]] <- ggplot(count(athletes, sex)) +
      geom_bar(aes(sex, n, fill = sex), stat = "identity") +
      scale_y_continuous(expand = c(0, 0)) +
      scale_fill_brewer(palette = "Set1") +
      labs(y = "number")
    
    p[["scatter"]] <- ggplot(athletes) +
      geom_point(aes(rcc, wcc, col = sex)) +
      scale_color_brewer(palette = "Set1") +
      theme(legend.position = "none") +
      labs(x = "RBC count", y = "WBC Count")
    
    p[["box"]] <- ggplot(athletes) +
      geom_boxplot(aes(sport, pcBfat, col = sex, fill = sex), alpha = 0.5) +
      scale_color_brewer(palette = "Set1") +
      scale_fill_brewer(palette = "Set1") +
      theme(legend.position = "none") +
      labs(y = "% body fat", x = NULL)
    ```
  
1. Now, we use `patchwork` to combine the subplots using the different
combinations discussed above.
    
    ```{r}
    p[["bar"]] + p[["scatter"]] + p[["box"]]
    p[["bar"]] / p[["scatter"]] / p[["box"]]
    (p[["bar"]] + p[["scatter"]]) / p[["box"]]
    ```

1. A corollary of using the same encodings across panels is that it should be
possible to share legends across the entire compound figure. This is most
concisely done by setting `plot_layout(legend = "collect")`. For example,
compare the athlete physiology dataset with and without the collected legends,
    
    ```{r}
    (p[["bar"]] + p[["scatter"]] + theme(legend.position = "left")) / p[["box"]] # turns legends back on
    ```

    The version with the legends collected is given below.

    ```{r}
    (p[["bar"]] + p[["scatter"]]) / p[["box"]] +
          plot_layout(guides = "collect") &
          plot_annotation(theme = theme(legend.position = "bottom"))
    ```
  
1. For annotation, we can add a title to each figure individually using
`ggtitle()`, before they are combined into the compound figure. The size and
font of the titles can be adjusted by using the `theme(title =
element_text(...))` option. For example, the code below adds the a - c titles
for each subpanel.
    
    ```{r}
    p[["bar"]] <- p[["bar"]] + ggtitle("a")
    p[["scatter"]] <- p[["scatter"]] + ggtitle("b")
    p[["box"]] <- p[["box"]] + ggtitle("c")
    
    (p[["bar"]] + p[["scatter"]]) / p[["box"]] +
      plot_layout(guides = "collect") &
      plot_annotation(theme = theme(legend.position = "bottom", title = element_text(size = 10)))
    ```
    
1. Patchwork handles alignment in the background, but sometimes we might want to
have control over the relative sizes of different panels. For this, we can again
use the `plot_layout` function, this time using the height and width arguments.
For example, the two examples change the height and widths of the first
component in the layout.

```{r}
    (p[["bar"]] + p[["scatter"]] + plot_layout(widths = c(1, 3))) / p[["box"]] +
      plot_layout(guides = "collect")

    (p[["bar"]] + p[["scatter"]]) / p[["box"]] +
      plot_layout(guides = "collect", heights = c(1, 3))
```

