---
title: "A Vocabulary of Marks"
description: |
  Examples of encodings and sequential refinement of a plot.
author:
  - name: Kris Sankaran
layout: post
output:
  distill::distill_article:
    self_contained: false
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE, echo = TRUE, fig.height = 5, retina = 4)
```

_[Reading](https://clauswilke.com/dataviz/aesthetic-mapping.html), [Recording](https://mediaspace.wisc.edu/media/Week%201%20%5B2%5D%20A%20Vocabulary%20of%20Marks/1_1tjacb3v),  [Rmarkdown](https://github.com/krisrs1128/stat436_s24/blob/main/notes/2024-12-27-week01-02.Rmd)_

1. The choice of encodings can have a strong effect on (1) the types of
comparisons that a visualization suggests and (2) the chance that readers leave with complete and accurate conclucions.
With this in mind, it's worthwhile
to develop a rich vocabulary of potential visual encodings. 

2. So, let’s look at a few different types of encodings available in `ggplot2`.
Before we get started, let's load up the libraries that will be used in these
notes. `ggplot2` is our plotting library. `readr` is used to read data files
from a web link, and `dplyr` is useful for some of the data manipulations below
(we dive into it deeply in Week 2).

```{r}
library(readr)
library(ggplot2)
library(dplyr)
theme_set(theme_bw()) # create a simpler default theme
```

### Point Marks

3. Let's read in the gapminder dataset, which describes the changes in standard
of living around the world over the last few decades. The `%>%` "pipe" operator
takes the output of the previous command as input to the next one -- it is
useful for chains of commands where the intermediate results are not needed. The
`mutate` command makes sure that the country group variable is treated as a
categorical, and not numeric, variable.

```{r}
gapminder <- read_csv("https://uwmadison.box.com/shared/static/dyz0qohqvgake2ghm4ngupbltkzpqb7t.csv", col_types = cols()) %>%
  mutate(cluster = as.factor(cluster))  # specify that cluster is nominal
gap2000 <- filter(gapminder, year == 2000) # keep only year 2000
```

4. Point marks can encode data fields using their $x$ and $y$ positions, color,
size, and shape. Below, each mark is a country, and we're using shape and the
$y$ position to distinguish between country clusters.

```{r}
ggplot(gap2000) +
  geom_point(aes(x = fertility, y = cluster, shape = cluster))
```

Since the first two arguments in `aes` are always the `x` and `y` positions, we
can omit it from our command. The code below produces the exact same plot (try
it!).

```{r, eval = FALSE}
ggplot(gap2000) +
  geom_point(aes(fertility, cluster, shape = cluster))
```

5. We can specify different types of shapes using the `shape` parameter outside of
the `aes` encoding.

```{r}
ggplot(gap2000) +
  geom_point(aes(fertility, cluster), shape = 15)
```

### Bar Marks

6. Bar marks let us associate a continuous field with a nominal one.

```{r}
ggplot(gap2000) +
  geom_bar(aes(country, pop), stat = "identity")
```

The plot above is messy -- it would not be appropriate for a publication or
presentation. The grid lines associated with each bar are distracting. Further,
the axis labels are all running over one another. For the first issue, we can
customize the `theme` of the plot. Note that we don't have to memorize the names
of these arguments, since they should autocomplete when pressing tab (we just
need to memorize the first few letters).

```{r}
ggplot(gap2000) +
  geom_bar(aes(country, pop), stat = "identity") +
  theme(panel.grid.major.x = element_blank())
```

For the second issue, one approach is to turn the labels on their side, again by
customizing the theme.

```{r}
ggplot(gap2000) +
  geom_bar(aes(country, pop), stat = "identity") +
  theme(
    axis.text.x = element_text(angle = 90),
    panel.grid.major.x = element_blank()
  )
```

An approach I like better is to turn the bars on their side. This way, readers
don't have to tilt their heads to read the country names.

```{r, fig.height = 6.7, fig.width = 4.5}
ggplot(gap2000) +
  geom_bar(aes(pop, country), stat = "identity") +
  theme(panel.grid.major.y = element_blank()) # note change from x to y
```

I'm also going to remove the small tick marks associated with every name, again
because it seems distracting.

```{r, fig.height = 6.7, fig.width = 4.5}
ggplot(gap2000) +
  geom_bar(aes(pop, country), stat = "identity") +
  theme(
    panel.grid.major.y = element_blank(),
    axis.ticks = element_blank() # remove tick marks
  )
```

7. To make comparisons between countries with similar populations easier, we can
order them by population (alphabetical ordering is not that meaningful). To
compare clusters, we can color in the bars.

    ```{r, fig.height = 6.7, fig.width = 4.5}
    ggplot(gap2000) +
      geom_bar(aes(pop, reorder(country, pop), fill = cluster), stat = "identity") +
      theme(
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank()
      )
    ```

8. We've been spending a lot of time on this plot. This is because I want to
emphasize that a visualization is not just something we can get just by memorizing some
magic (programming) incantation. Instead, it is something worth critically
engaging with and refining, in a similar way that we would refine an essay or
speech.

    Philosophy aside, there are still a few points that need to be improved in this figure,
    
    * The axis titles are not meaningful.
    * There is a strange gap between the left hand edge of the plot and the start of the bars.
    * I would also prefer if the bars were exactly touching one another, without the small vertical gap.
    * The scientific notation for population size is unnecessarily technical.
    * The color scheme is a bit boring.
      
    I'll address each of these in a separate code block, with comments on the
    parts that are different. First, improving the axis titles,
        
    ```{r, fig.height = 6.7, fig.width = 4.5}
    ggplot(gap2000) +
      geom_bar(aes(pop, reorder(country, pop), fill = cluster), stat = "identity") +
      labs(x = "Population", y = "Country", fill = "Country Group") + # add better titles
      theme(
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank()
      )
    ```
      
    Now we remove the gap. I learned this trick by [googling
    it](https://stackoverflow.com/questions/22945651/remove-space-between-plotted-data-and-the-axes)
    -- there is no shame in doing this! A wise friend of mine once shared, "I am
    not a programming expert, just an expert at StackOverflow."
    
    ```{r, fig.height = 6.7, fig.width = 4.5}
    ggplot(gap2000) +
      geom_bar(aes(pop, reorder(country, pop), fill = cluster), stat = "identity") +
      scale_x_continuous(expand = c(0, 0, 0.1, 0.1)) + # remove space to the axis
      labs(x = "Population", y = "Country", fill = "Country Group") + 
      theme(
        axis.text.y = element_text(size = 6),
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank()
      )
    ```

    Now, removing the gaps between bars. 
    ```{r, fig.height = 6.7, fig.width = 4.5}
    ggplot(gap2000) +
      geom_bar(
        aes(pop, reorder(country, pop), fill = cluster),
        width = 1, stat = "identity" # increase width of bars
      ) +
      scale_x_continuous(expand = c(0, 0, 0.1, 0.1)) +
      labs(x = "Population", y = "Country", fill = "Country Group", color = "Country Group") +
      theme(
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank()
      )
    ```
    
    Now, we remove scientific notation,
    ```{r, fig.height = 6.7, fig.width = 4.5}
    ggplot(gap2000) +
      geom_bar(
        aes(pop, reorder(country, pop), fill = cluster),
        width = 1, stat = "identity"
      ) +
      scale_x_continuous(label = scales::label_number_si(), expand = c(0, 0, 0.1, 0.1)) + # remove scientific notation. scales::comma() is also useful.
      labs(x = "Population", y = "Country", fill = "Country Group", color = "Country Group") +
      theme(
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank()
      )
    ```
    
    Finally, we customize the colors. I often like to look up neat colors on
    [color.adobe.com](https://color.adobe.com/),
    [iwanthue](https://medialab.github.io/iwanthue/) or
    [colorhexa](https://www.colorhexa.com/), but there are dozens of similar
    colorpicker sites out there.
    
    ```{r, fig.height = 6.7, fig.width = 4.5}
    ggplot(gap2000) +
      geom_bar(
        aes(pop, reorder(country, pop), fill = cluster),
        width = 1, stat = "identity"
      ) +
      scale_x_continuous(label = scales::label_number_si(), expand = c(0, 0, 0.1, 0.1)) +
      scale_fill_manual(values = c("#80BFA2", "#7EB6D9", "#3E428C", "#D98BB6", "#BF2E21", "#F23A29")) +
      labs(x = "Population", y = "Country", fill = "Country Group", color = "Country Group") +
      theme(
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank()
      )
    ```

9. This seems like a lot of work for just a lowly bar plot! But I think it's
amazing customizable the figure is -- we can give it our own sense of style.
With a bit of practice, these sorts of modifications will become second nature,
and it won't be necessary to keep track of all the intermediate code. And
really, even though we spent some time on this plot, there are still many things
that could be interesting to experiment with, like font styles, background
appearance, maybe even splitting the countries into two panels.

10. In the plot above, each bar is anchored at 0. Instead, we could have each
bar encode *two* continuous values, a left and right. To illustrate, let's
compare the minimum and maximimum life expectancies within each country cluster.
We'll need to create a new `data.frame` with just the summary information. For
this, we `group_by` each cluster, so that a `summarise` call finds the minimum
and maximum life expectancies restricted to each cluster. We'll discuss the
`group_by` + `summarise` pattern in detail next week.
    
    ```{r, fig.height = 4.5}
    # find summary statistics
    life_ranges <- gap2000 %>%
      group_by(cluster) %>%
      summarise(
        min_life = min(life_expect),
        max_life = max(life_expect)
      )
    
    # look at a few rows
    head(life_ranges)
    ggplot(life_ranges) +
      geom_segment(
        aes(min_life, reorder(cluster, max_life), xend = max_life, yend = cluster, col = cluster),
        size = 5,
      ) +
      scale_color_manual(values = c("#80BFA2", "#7EB6D9", "#3E428C", "#D98BB6", "#BF2E21", "#F23A29")) +
      labs(x = "Minimum and Maximum Expected Span", col = "Country Group", y = "Country Group") +
      xlim(0, 85) # otherwise would only range from 42 to 82
    ```
    
### Line Marks

11. Line marks are useful for comparing changes. Our eyes naturally focus on
rates of change when we see lines. Below, we'll plot the fertility over time,
colored in by country cluster. The `group` argument is useful for ensuring each
country gets its own line; if we removed it, `ggplot2` would become confused by
the fact that the same `x` (year) values are associated with multiple `y`'s
(fertility rates).
    
    ```{r}
    ggplot(gapminder) +
      geom_line(
        aes(year, fertility, col = cluster, group = country),
          alpha = 0.7, size = 0.9
      ) +
      scale_x_continuous(expand = c(0, 0)) +  # same trick of removing gap
      scale_color_manual(values = c("#80BFA2", "#7EB6D9", "#3E428C", "#D98BB6", "#BF2E21", "#F23A29"))
    ```
    
### Area Marks

12. Area marks have a flavor of both bar and line marks. The filled area supports
absolute comparisons, while the changes in shape suggest derivatives.
    
    ```{r}
    population_sums <- gapminder %>%
      group_by(year, cluster) %>%
      summarise(total_pop = sum(pop))
    head(population_sums)
    
    ggplot(population_sums) +
      geom_area(aes(year, total_pop, fill = cluster)) +
      scale_y_continuous(expand = c(0, 0, .1, .1), label = scales::label_number_si()) +
      scale_x_continuous(expand = c(0, 0)) +
      scale_fill_manual(values = c("#80BFA2", "#7EB6D9", "#3E428C", "#D98BB6", "#BF2E21", "#F23A29"))
    ```
    
    Just like in bar marks, we don't necessarily need to anchor the $y$-axis at 0.
    For example, here the bottom and top of each area mark is given by the 30% and
    70% quantiles of population within each country cluster.
    
    ```{r}
    population_ranges <- gapminder %>%
      group_by(year, cluster) %>%
      summarise(min_pop = quantile(pop, 0.3), max_pop = quantile(pop, 0.7))
    head(population_ranges)
    
    ggplot(population_ranges) +
      geom_ribbon(
        aes(x = year, ymin = min_pop, ymax = max_pop, fill = cluster),
        alpha = 0.8
      ) +
      scale_y_continuous(expand = c(0, 0, .1, .1), label = scales::label_number_si()) +
      scale_x_continuous(expand = c(0, 0)) +
      scale_fill_manual(values = c("#80BFA2", "#7EB6D9", "#3E428C", "#D98BB6", "#BF2E21", "#F23A29"))
    ```
    
