---
title: "Linking using Crosstalk"
description: |
  Linking in web-based visualizations.
author:
  - name: Kris Sankaran
layout: post
output:
  distill::distill_article:
    self_contained: false
always_allow_html: true
---

```{r, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(cache = FALSE, message = FALSE, warning = FALSE, echo = TRUE, fig.height = 5, retina = 4)
```

_[Reading](https://emilyriederer.netlify.app/post/crosstalk/), [Recording](https://mediaspace.wisc.edu/media/t/1_h7dh543g),  [Rmarkdown](https://github.com/krisrs1128/stat436_s24/blob/main/notes/2024-12-27-week05-04.Rmd)_

```{r}
library(ggplot2)
library(plotly)
library(leaflet)
library(DT)
library(crosstalk)
```

1. For most graphical queries, the click and brush inputs implemented in Shiny
will be sufficient. However, a basic limitation of shiny's `plotOutput` is that
it has to create views by generating static image files -- it only creates the
illusion of interactivity by rapidly changing the underlying files. In some
cases, there will be so many points on the display each update will be slow and
the fluidity of interaction will suffer.

1. One approach around this problem is to use a library that directly supports
web-based plots. These plots can modify elements in place, without having to
redraw and save the entire figure on each interaction. The `crosstalk` package
gives one approach to linked views in this setting. We'll only give an overview
of it here, but the purpose of sharing it is so we have at least one example
that we can refer to in case the Shiny approach becomes untenable.

1. We'll study a problem about the dropoff in Chicago subway ridership after the
start of the COVID-19 lockdowns. We have data on the weekday and weekend transit
ridership at each subway station, along with the locations of the stations. We
are curious about the extent of the change in ridership, along with features
that might be responsible for some stations being differentially affected. The
block below reads in this raw data,
    
    ```{r}
    download.file("https://github.com/emilyriederer/demo-crosstalk/blob/master/data/stations.rds?raw=true", "stations.rds")
    download.file("https://github.com/emilyriederer/demo-crosstalk/blob/master/data/trips_apr.rds?raw=true", "trips_apr.rds")
    stations <- readRDS("stations.rds")
    trips <- readRDS("trips_apr.rds")
    ```
    
1. The crosstalk package implements a `SharedData` object. This is used to track
selections across all the plots that refer to it. We can think of it as
crosstalk's analog of our earlier `brushedPoints` function. These objects are
defined by calling `SharedData$new()` on the data.frame which will be used
across views. The `key` argument provides a unique identifier that is used to
match corresponding samples across all displays (notice that it uses `~` formula
notation).

    ```{r}
    trips_ct <- SharedData$new(trips, key = ~station_id, group = "stations")
    ```
    
1. Let's see how this object can be used for linked brushing. We'll first
generate static ggplot2 objects giving (1) a view of weekday ridership in 2019
vs. 2020 and (2) a view of what proportion of 2019 ridership at the station took
place on weekends.
    
    ```{r, eval = FALSE}
    p1 <- ggplot(trips_ct) +
      geom_point(aes(year_2019_wday, year_2020_wday)) +
      geom_abline(slope = 1, col = "#0c0c0c")
    
    p2 <- ggplot(trips_ct) +
      geom_point(aes(prop_wend_2019, reorder(station_name, prop_wend_2019)), stat = "identity")
    ```
  
1. Given this ggplot2 base, we can build web-based `plotly` objects using the
`ggplotly` command. The `layout` and `highlight` functions are specifying that
we want user interactions to define brushes, not zoom events. The final `bscols`
function allows us to place the views side by side. By brushing the two plots,
we can see that those stations with the largest dropoff in riderships were those
that were mostly used during the weekdays. This makes sense, considering many of
the office workers in Chicago started working from home in 2020.
    
    ```{r, eval = FALSE}
    p1 <- ggplotly(p1, tooltip = "station_name") %>%
      layout(dragmode = "select") %>%
      highlight(on = "plotly_selected")
    
    p2 <- ggplotly(p2) %>%
      layout(dragmode = "select", direction = "v") %>%
      highlight(on = "plotly_selected")
    
    bscols(p1, p2)
    ```

1. This `crosstalk` approach works with more than plotly-derived plots. In the
block below, we also generate a map (using the `leaflet` package) and a data
table (using the `DT` package). The views are all synchronized because they
refer to the same `station_id` key in `SharedData` objects. This visualization
confirms our intuition that those stations with the largest drop-off in
ridership are those that are downtown.

```{r, eval = FALSE}
stations_ct <- SharedData$new(stations, key = ~station_id, group = "stations")
dt <- datatable(trips_ct)
lf <- leaflet(stations_ct) %>% 
  addTiles() %>% 
  addMarkers()

bscols(p1, p2, lf, dt, widths = rep(6, 4))
```
