---
title: "Seasonal Plots"
description: |
  Approaches for visualizing seasonality.
author:
  - name: Kris Sankaran
    affiliation: UW Madison
layout: post
output:
  distill::distill_article:
    self_contained: false
---

_[Reading](https://otexts.com/fpp3/seasonal-plots.html), [Recording](https://mediaspace.wisc.edu/media/Week%206%20%5B3%5D%20Seasonal%20Plots/1_vbhlrnx5), [Rmarkdown](https://github.com/krisrs1128/stat479/blob/master/_posts/2021-02-23-week6-3/week6-3.Rmd)_

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, message = FALSE, warning = FALSE, echo = TRUE)
```

```{r}
library(tidyverse)
library(feasts)
library(fpp2)
library(tsibbledata)
theme_set(theme_minimal())
```


1. If our data have seasonal structure, it’s natural to compare individual
periods against one another. In contrast to plotting the data in one long time
series, seasonal plots reduce the amount of distance our eyes have to travel in
order to compare two periods in a seasonal pattern. This also reduces the burden
on our memory.

2. In R, we can use the `gg_season` function to overlay all the seasons onto one
another. The plot below shows antidiabetes drug sales over time. This view makes
it clear that there is a spike in sales every January.

```{r}
cols <- scales::viridis_pal()(10)
gg_season(as_tsibble(a10), pal = cols)
```

3. If the time series exhibit seasonal structure at multiple scales, then
we can view them all using the `period` argument.

```{r}
gg_season(vic_elec, Demand, period = "day", pal = cols)
gg_season(vic_elec, Demand, period = "week", pal = cols)
```
