---
title: "Spatial Data Formats"
description: |
  An overview of common formats, with illustrative examples.
author:
  - name: Kris Sankaran
    affiliation: UW Madison
date: 2024-01-10
layout: post
output:
  distill::distill_article:
    self_contained: false
---

_[Reading](https://geocompr.robinlovelace.net/spatial-class.html), [Recording](https://mediaspace.wisc.edu/media/Week%207%20%5B1%5D%20Spatial%20Data%20Formats/1_r6gnm2cv), [Rmarkdown](https://github.com/krisrs1128/stat436_s24/blob/main/notes/2024-12-27-week07-01.Rmd)_

```{r setup}
knitr::opts_chunk$set(cache = FALSE, message = FALSE, warning = FALSE, echo = TRUE, eval = TRUE)
```

```{r}
library(ceramic)
library(raster)
library(sf)
library(terra)
library(tidyverse)
library(tmap)
theme_set(theme_minimal())
```

1. Spatial data come in two main formats: vector and raster. We’ll examine them
in detail in the next few lectures, but this lecture motivates the high-level
distinction and gives a few examples. It also shows how to read and write data
to and from these formats.

### Vector Data

2. Vector data formats are used to store geometric information, like the
locations of hospitals (points), trajectories of bus routes (lines), or
boundaries of counties (polygons). It’s useful to think of the associated data
as being spatially enriched data frames, with each row corresponding to one of
these geometric features.

3. Vector data are usually stored in `.geojson`, `.wkt`, `.shp`, or `.topojson`
formats. Standard data.frames cannot be used because then important spatial
metadata would be lost, like the Coordinate Reference System (to be explained in
the fourth lecture this week).

4. In R, these formats can be read using `read_sf` in the `sf` package. They can
be written using the `write_sf` function. Here, we'll read in a vector dataset
containing the boundaries of lakes in Madison. 

    ```{r}
    lakes <- read_sf("https://uwmadison.box.com/shared/static/duqpj0dl3miltku1676es64d5zmygy92.geojson")
    
    lakes %>%
      dplyr::select(id, name, geometry)
    #write_sf(lakes, "output.geojson", driver = "GeoJSON")
    ```

    We'll discuss plotting in the next lecture, but for a preview, this is how you
    can visualize the lakes using ggplot2.
  
    ```{r, echo = TRUE}
    lakes <- lakes %>%
      group_by(id) %>%
      mutate(
        longitude = st_coordinates(geometry)[1, 1],
        latitude = st_coordinates(geometry)[1, 2]
      )
    
    tm_shape(lakes) +
      tm_polygons(col = "#00ced1")
    ```
  
    With a little extra effort, we can overlay the features onto public map
    backgrounds (these are often called "basemaps").

    ```{r, echo = TRUE}
    # you can get your own at https://account.mapbox.com/access-tokens/
    Sys.setenv(MAPBOX_API_KEY="pk.eyJ1Ijoia3Jpc3JzMTEyOCIsImEiOiJjbDYzdjJzczQya3JzM2Jtb2E0NWU1a3B3In0.Mk4-pmKi_klg3EKfTw-JbQ")
    basemap <- cc_location(loc= c(-89.401230, 43.073051), buffer = 15e3)
    
    tm_shape(basemap) +
      tm_rgb() +
      tm_shape(lakes) +
      tm_polygons(col = "#00ced1")
    ```
    
5. There is a surprising amount of public vector data available online. Using this [query^[It can be constructed easily using the wizard]](https://overpass-turbo.eu/s/14ma), I’ve downloaded locations of all hospital clinics in Madison.

    ```{r}
    clinics <- read_sf("https://uwmadison.box.com/shared/static/896jdml9mfnmza3vf8bh221h9hlvh70v.geojson")
    
    # how would you overlay the names of the clinics, using geom_text?
    tm_shape(basemap) +
      tm_rgb() +
      tm_shape(clinics) +
      tm_dots(col = "red", size = 1)
    ```
  	
    Using this [query](https://overpass-turbo.eu/s/14m9), I’ve downloaded all the
    bus routes.

    ```{r}
    bus <- read_sf("https://uwmadison.box.com/shared/static/5neu1mpuh8esmb1q3j9celu73jy1rj2i.geojson")
    
    tm_shape(basemap) +
      tm_rgb() +
      tm_shape(bus) +
      tm_lines(col = "#bc7ab3", size = 1)
    ```
    
    For the boundaries of the lakes above, I used this
    [query](https://overpass-turbo.eu/s/14mb).
    	
    Many organizations prepare geojson data themselves and make it publicly
    available; e.g., the boundaries of
    [rivers](https://data-wi-dnr.opendata.arcgis.com/datasets/c7de80dd473d440e98ab3acb611e7d64_2)
    or
    [glaciers](https://rds.icimod.org/Home/DataDetail?metadataId=9361&searchlist=True).
    Don’t worry about how to visualize these data at this point — I just want to
    give some motivating examples.
    
### Raster Data

6. Raster data give a measurement along a spatial grid. You can think of them as
spatially enriched matrices, where the metadata says where on the earth each
entry of the matrix is associated with.

7. Raster data are often stored in `tiff` format. They can be read in using the
`rast` function in the `terra` library, and can be written using
`writeRaster`.

    ```{r}
    shanghai <- rast("https://uwmadison.box.com/shared/static/u4na56w3r4eqg232k2ma3eqbvehfiaoq.tif")
    #writeRaster(shanghai, "output.tiff", driver = "GeoTIFF")
    ```
  
8. Some of the most common types of public raster data are satellite images or
derived measurements, like elevation maps. For example, the code below shows an
image of a neighborhood outside Shanghai.

```{r}
tm_shape(shanghai / 1636 * 255) + # rescale the max value to 255
  tm_rgb()
```

There’s actually quite a bit of information in this image. We can zoom in…

```{r}
bbox <- ext(121.66, 121.665, 30.963, 30.968)
shanghai_ <- crop(shanghai, bbox)

tm_shape(shanghai_ / 1636 * 255) +
  tm_rgb()
```

Here are is data on elevation in Zion national park.

```{r}
f <- system.file("raster/srtm.tif", package = "spDataLarge")
zion <- rast(f)
tm_shape(zion) +
  tm_raster(palette = "PuBuGn") +
  tm_layout(legend.position = c("left", "bottom"))
```
  
### Installation

9. A note about R packages: for historical reasons, spatial data libraries in R
reference a few command line programs, like `gdal` and `proj`. Since these
command line programs are not themselves a part of R, they need to be installed
before the corresponding R packages. The process will differ from operating
system to operating system, and the experience can be frustrating, especially
when the R packages don’t recognize the underlying system installation. I
recommend following the instructions on [this
page](https://github.com/r-spatial/sf) and reaching out early if you have any
issues.
